Relevance Modeling with Documents Expanded Using External Collections

### 2017-02-26 expansion/paper.tex
added QPP paper with initial work
page limit and small fixes

\author{Garrick Sherman}

@@

\begin{flalign}\label{equation.ql}
\log P(Q | D) = \prod_{w \in Q} P(w | Q) \cdot \log P(w | \theta_D) .
\end{flalign}

\noindent We follow standard procedures for estimating the probabilities in Eq. \ref{equation.ql}.  We simply use the maximum likelihood estimate of $\hat{P}(w | Q) = \frac{c(w, Q)}{|Q|}$ where $c(w, Q)$ is the frequency of word $w$ in $Q$.  For $P(w | \theta_D)$ we estimate a smoothed language model by assuming that document language models
in a given collection have a Dirichlet prior distribution:

@@

A relevance model takes the form of

\begin{flalign}\label{equation.rm1}
	P(w|R) = \sum_{D \in C} P(D) P(w|D) P(Q|D)
\end{flalign}

\noindent where $P(Q|D)$ is calculated as in Eq. \ref{equation.ql} and essentially weights word $w$ in $D$ by the query likelihood of the document. Relevance models are most efficient and robust when calculated over only the top terms in only the top ranked documents. These parameters are referred to as $fbTerms$ and $fbDocs$ respectively in Table \ref{table.parameters}.

@@

We obtain a ranking of related documents, which we call expansion documents, by running $Q_D$ over an index $C_E$. More formally, we rank the documents in $C_E$ against $D$ using Eq. \ref{equation.ql}, substituting $Q_D$ for the query and $E_i$---the text of the $i^{th}$ expansion document---for the document. Let $\pi_i$ be the log-probability for expansion document $E_i$ with respect to $D$ given by Eq. \ref{equation.ql}.

@@

\subsubsection{Expansion Source}\label{section.evaluation.runs.expansion}

We test each type of expansion source individually, with documents expanded using:
\begin{itemize}
	\item no expansion, called \textit{baseline};
	\item the target collection itself, called \textit{self};
	\item Wikipedia, called \textit{wiki}; or
	\item a mixture of the previous two, called \textit{combined}.
\end{itemize}

@@

Figure \ref{figure.n-sweeps} shows sweeps over several values of $n$, the number of expansion documents, for the \textit{self} and \textit{wiki} QL runs using our established cross-validation procedure with identical folds. The sensitivity to $n$ is not pronounced at $n \geq 5$, and what little variation exists is not consistent across collections. We therefore set the value of 10, an apparently safe value, for all other runs. This is a convenient result since it allows for more efficient document expansion.

@@

The results indicate that our approach for document expansion work well in general and especially in concert with traditional relevance modeling techniques. We find that we can improve on traditional document expansion by incorporating external collections into the expansion process. In the future, we plan to investigate in detail the relationship between homogeneity of documents and the effectiveness of our model.

\section{Acknowledgments}\label{section.acknowledgments}
This work was supported in part by the US National Science Foundation under Grant No. [blind]. Any opinions, findings, conclusions, or recommendations expressed are those of the authors and do not necessarily reflect the views of the National Science Foundation.

>>>>

%\author{Garrick Sherman}
\author{[Blind]}

@@

\begin{flalign}\label{equation.ql}
P(Q | D) = \prod_{w \in Q} P(w | \theta_D)^{c(w,Q)}
\end{flalign}

\noindent where $c(w, Q)$ is the frequency of word $w$ in $Q$. We follow standard procedures for estimating $P(w|\theta_D)$ in Eq. \ref{equation.ql}, estimating a smoothed language model by assuming that document language models in a given collection have a Dirichlet prior distribution:

@@

Assuming uniform document prior probabilities, relevance models take the form of

\begin{flalign}\label{equation.rm1}
	P(w|R) = \sum_{D \in C} P(w|D) P(Q|D)
\end{flalign}

\noindent where $P(Q|D)$ is calculated as in Eq. \ref{equation.ql} and essentially weights word $w$ in $D$ by the query likelihood of the document. Relevance models are most efficient and robust when calculated over only the top ranked documents and limited to the top terms. These parameters are referred to as $fbDocs$ and $fbTerms$ respectively in Table \ref{table.parameters}.

@@

We rank related documents, called expansion documents, by running $Q_D$ over an index $C_E$. More formally, we rank the documents in $C_E$ against $D$ using Eq. \ref{equation.ql}, substituting $Q_D$ for the query and $E_i$---the $i^{th}$ expansion document---for the document. Let $\pi_i$ be the log-probability for expansion document $E_i$ with respect to $D$ given by Eq. \ref{equation.ql}.

@@

We produce runs with expansion documents from:
\begin{itemize}
	\item the target collection itself, called \textit{self};
	\item Wikipedia, called \textit{wiki};
	\item a mixture of the previous two, called \textit{combined}; or
	\item no expansion, called \textit{baseline};
\end{itemize}

@@

Figure \ref{figure.n-sweeps} shows sweeps over several values of $n$, the number of expansion documents, for the \textit{self} and \textit{wiki} QL runs using our established cross-validation procedure with identical folds. The sensitivity to $n$ is not pronounced at $n \geq 5$, and what little variation exists is not consistent across collections. We therefore set $n$ to 10, an apparently safe value, for all other runs. This is a convenient result since it allows for more efficient document expansion.

@@

The results indicate that our approach for document expansion works well in general and especially in concert with traditional relevance modeling. We find that we can improve on traditional document expansion by incorporating external collections into the expansion process. In the future, we plan to investigate in detail the relationship between homogeneity of documents and the effectiveness of our model.

%\section{Acknowledgments}\label{section.acknowledgments}

%This work was supported in part by the US National Science Foundation under Grant No. [blind]. Any opinions, findings, conclusions, or recommendations expressed are those of the authors and do not necessarily reflect the views of the National Science Foundation.
